DRAFT: Harmonisation and trait aggregation paper

### 2020-12-19 Paper/Draft_harmonization_paper.tex
19.12.: Started fixing references, fixed figure 7

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{float}
\usepackage[table,xcdraw]{xcolor}
\usepackage{color, colortbl}
\usepackage{longtable}
%\usepackage[sort&compress,square,comma,authoryear]{natbib}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{rotating}
% \graphicspath{{/home/kunz/Dokumente/Projects/Trait_DB/Invertebrate_traits/Paper/Figures/}}
\graphicspath{{Figures/}}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{array}
\usepackage{lscape}
\usepackage{longtable}
\usepackage{hhline}
\usepackage{lmodern}
\usepackage[
backend = biber,
style = ieee,
citestyle = numeric-comp,
maxbibnames=99,
sorting = none % sort by name year title
]{biblatex}
\addbibresource{Ref_invertebrate_DB.bib}

@@

Explaining and predicting how communities are shaped by environmental factors is one of the main goals of ecology. Species traits, defined as measurable properties of an organism \cite{mcgill_rebuilding_2006}, might be beneficial in achieving this goal \cite{heino_jani_macroecological_2013}. Traits evolve through adaptations (e.g., physiological, behavioural, etc.) of organisms to their environment and indicate direct or indirect linkages between the biological response of an organism or a population and its environment \cite{southwood_habitat_1977, verberk_delivering_2013}.
Besides providing a mechanistic explanation of species-environment relationships, trait-based approaches might be suitable for large scale analysis because the variability in trait responses is lower than in taxonomic responses \cite{bonada_taxonomic_2007, baird_toward_2011}. Traits of freshwater invertebrate individuals are difficult to determine because - unlike plant traits - they often can not be measured directly. For example, to gain knowledge on feeding habits requires evaluating mouthpart morphology, consumed food, and the organisms function within its community \cite{moog_comprehensive_nodate}. Nevertheless, invertebrate traits have been increasingly used in freshwater ecology, e.g. by relating macroinvertebrate trait composition to environmental factors and as trait metrics in biomonitoring \cite{poff_developing_2010, szocs_effects_2014, bhowmik_large_2015, menezes_beyond_2010}.

In the last decades, freshwater ecologists have compiled comprehensive invertebrate trait databases for various continents \cite{usseglio-polatera_biomonitoring_2000, schmidt-kloiber_www.freshwaterecology.info_2015, vieira_database_nodate, Philips_and_Smith_NZ_DB_2018, kefford_integrated_2020, tomanova_trophic_2006}. The availability of invertebrate trait data from different continents enables comparisons of trait variation and analyses of their relation to environmental factors across large scales. However, such analyses have been carried out mostly within continents, using information from one or two trait databases. For example, Bonada et al. \cite{bonada_taxonomic_2007} compared trait composition for Mediterranean and temperate regions in Europe using traits from Usseglio-Polatera et al. \cite{usseglio-polatera_biomonitoring_2000}(typically referred to as Tachet database); Poff et al. \cite{poff_developing_2010} characterized trait composition across sites in the Western US using traits from Poff et al. \cite{poff_functional_2006}, and Botwe et al. \cite{botwe_effects_2018} investigated the effect of salinity on invertebrate traits in different sites in South Australia using trait data from Poff et al. \cite{poff_functional_2006} and Schäfer et al. \cite{schafer_trait_2011}. Analyses of invertebrate traits that synthesize information on invertebrate grouping features from more than two different continents are rare.

In this study, we follow the terminology proposed by Schmera et al. \cite{schmera_proposed_2015}, where a grouping feature is defined as a general property (e.g. feeding mode) that comprises a "group of related traits (e.g., predator, shredder, etc.) that vary among species or among individuals within a species". Thus, we use the term grouping feature in place of which many studies use the term "trait", and the term trait instead of "trait state", "modality" or "trait category". Traits can be described using different codings (binary, fuzzy) that represent the prevalence of a characteristic in an organism.

To our knowledge, only Brown et al. \cite{brown_functional_2018} harmonised grouping features from more than two geographically distant invertebrate trait databases, for a limited set of grouping features (8) and taxa (112), in a study on the influence of decreasing glacier cover on functional diversity and community assembly of invertebrates. We suspect that the heterogeneity of information in freshwater invertebrate trait databases, besides the diversity of taxa across regions, is likely a major reason for the lack of studies across continents. To harmonize grouping features from different regions, first commonly accepted and unambiguous trait definitions are required \cite{schneider_towards_2019}. In the best case, grouping features would be classified into the same traits across databases or they could easily be harmonised using standardised terminology. However, a lack of standardised terminology of trait definitions and poor metadata quality in many trait databases are common issues throughout the field of trait-based ecology \cite{baird_toward_2011, kissling_towards_2018}. Secondly, consistent coding of traits facilitates the compatibility of trait data from different databases.

@@

However, invertebrate trait databases are heterogeneous regarding the coding they use for their traits \cite{culp_incorporating_2011}. For example, in the study from Brown et al. \cite{brown_functional_2018} the authors needed to reclassify traits using expert knowledge given that the individual databases employed different coding approaches (i.e. European and New Zealand databases fuzzy coding, North American database binary coding). Databases with standardised and unambiguous traits and a consistent coding of traits would minimize the data processing effort. Discrepancies in the taxonomic resolutions (e.g. species, genus or family) when linking observational taxonomic data to trait database represents another challenge. When observations are on a more precise taxonomic level than data available in the trait databases (e.g. observations on species-level, trait data on genus-level) trait information of the less precise taxonomic level is often assigned (e.g. \cite{szocs_effects_2014, vos_taxonomic_2017}). Conversely, if trait information is only available on more precise taxonomic levels than the observed taxa, traits are aggregated to a less precise taxonomic level, e.g. \cite{poff_functional_2006, szocs_effects_2014, piliere_a._f._h._importance_2016, aspin_extreme_2019}. Studies have used different methods of trait aggregation, e.g. the mean \cite{magliozzi_functional_2019}, median \cite{szocs_effects_2014} or the mode \cite{piliere_a._f._h._importance_2016}. Up to now, studies on how and to which extent different trait aggregation methods influence trait-based analysis are missing. However, related knowledge would inform future studies regarding the choice of the aggregation method.

The outlined difficulties when working with or synthesizing trait data motivated us to (1) investigate discrepancies in trait definition between trait databases from Europe, North America, Australia, and New Zealand. As a basis for further evaluating different trait aggregation methods and how harmonising definition discrepancies between grouping features can influence trait-environment relationships, we established 4 novel invertebrate trait datasets from these trait databases by harmonising 7 grouping features. We evaluate different trait aggregation methods by (2) comparing aggregated traits with assigned traits by experts at family-level for two of the established trait datasets. (3) we re-analysed data on the effect of anthropogenic salinisation on biological traits by Szöcs et al. \cite{szocs_effects_2014} using trait data from the established harmonised European trait datasets and aggregated traits. By comparison with the original analysis, we investigated how harmonising and aggregating trait data can modify the outcome of trait-environment relationships.

@@

Using harmonised grouping features resulted in less traits that distinguish upstream and downstream sites in comparison to the original analysis (Figure \ref{fig:rda_species_scores} and Figure \ref{fig:boxplots_scores_on_constrained_axis}). According to the RDA of the trait composition, downstream sites were characterised by taxa with the traits multivoltinism, ovoviviparity and long life cycle duration. Upstream sites were characterised by univoltine taxa with a short life cycle duration that lay their eggs in an aquatic environment (aquatic eggs). The trait aquatic eggs of the harmonised grouping feature oviposition has been derived by amalgamating the trait oviposition in clutches and other related traits (Table \ref{tab:traits_harmonisation}). The traits shredder and gill respiration did not characterise sites with high salinisation.

Using at family-level aggregated traits from the harmonised dataset showed also results similar to the original analysis but with fewer traits distinguishing upstream and downstream sites (Figure \ref{fig:rda_species_scores} and Figure \ref{fig:boxplots_scores_on_constrained_axis}). The \textit{direct\_agg\textsubscript{mean}}, \textit{direct\_agg\textsubscript{median}}, and \textit{weighted\_agg} 
characterised the downstream sites with the same traits as the original analysis except that downstream sites were not characterised by the trait shredder. Upstream sites were characterised by the traits univoltinism, aquatic eggs and short life cycle duration. The same results were obtained when re-analysing the data with traits aggregated by the \textit{stepwise\_agg\textsubscript{mean}} and \textit{stepwise\_agg\textsubscript{median}}, with the exception that none of the life cycle traits characterised upstream or downstream sites. Thus, in our comparison aggregating by \textit{direct\_agg\textsubscript{mean}}, \textit{direct\_agg\textsubscript{median}}, and \textit{weighted\_agg} yielded to the least change in interpretation of the RDA results compared to the original results.

\begin{figure}[H]
    \label{fig:rda_species_scores}
    \centering
    \includegraphics[width=16.5cm, height=10cm]{Species_scores_rda.png}
    \caption{Species scores obtained by RDA from the original analysis \cite{szocs_effects_2014}, using harmonised grouping features, and using harmonised grouping features with traits aggregated to family-level. The dots represent the individual species scores for each analysed trait along the conductivity axis. The violin plot shows the density estimate of the species scores. Gray vertical lines indicate the median of the obtained species scores.}
\end{figure}

% \begin{sidewaysfigure}
%   \label{fig:boxplots_scores_on_constrained_axis}
%   \centering
%   \includegraphics[width= 22cm, height=12cm]{Scores_on_constrained_axis_combined.pdf}
%   \caption{RDA of traits constrained by electric conductivity for the tested methods and the original study. Shown are boxplots of the site scores along the conductivity axis. The rug on the right side of each plot indicates species scores of the traits on the conductivity axis. Only traits with a mahalanobis distance greater than the $97.5 \%$ quantile of the Chi-square distribution (5.02) were labelled.}
% \end{sidewaysfigure}

\begin{figure}[H]
  \label{}
  \centering
  \includegraphics[width=16.5cm, height=25cm]{Scores_on_constrained_axis_comb_ALT.pdf}
  \caption{}
\end{figure}

@@

@@

Synthesizing trait information from multiple trait databases is crucial to unfold the full potential of trait-based approaches, e.g. for studying community trait responses to environmental gradients worldwide. Our attempt of harmonising invertebrate trait information from trait databases of different regions showed that grouping feature harmonisation is labour-intensive because grouping features are differentiated differently into traits between trait databases and various codings are used to describe traits. In addition, the same traits are sometimes defined differently in the investigated databases, requiring expert knowledge to reclassify these traits (e.g. the trait piercer in the Tachet database, Table \ref{tab:traits_harmonisation}). Although, others have noted the lack of standardised trait terminology in freshwater ecology \cite{baird_toward_2011, brink_traits-based_2011}, we are the first to describe invertebrate trait definition discrepancies for some commonly used grouping features.
%In addition we provide harmonised grouping feature trait datasets that can be used for other studies
To resolve definition discrepancies and facilitate data synthesis in the future, terminological standards are needed. Harmonised definitions and concepts of traits have been developed in the past for other organism groups, e.g. for plants with the \textit{Thesaurus Of Plant characteristics} (TOP) initiative \cite{garnier_towards_2017}. The core of this harmonisation initiative is to provide standardised trait definitions and to draw connections to synonyms, related terms, and surrounding concepts by linking to other controlled vocabularies or ontologies in the field. Such initiatives could be a role model for freshwater ecologists to establish unambiguous terminologies for invertebrate traits. The existing freshwater invertebrate trait databases could be linked through standardised terminologies or ontologies, as suggested by Baird et al. \cite{baird_toward_2011} almost a decade ago. By following the recently proposed \textit{Ecological Trait-data Standard Vocabulary} (ETS) providers of invertebrate trait data could connect their traits to such ontologies \cite{schneider_towards_2019}. Once standardised trait definitions are established, these will improve trait data sharing, trait data processing when working with multiple trait databases, and ultimately interpretation of the derived results. % What is not mentioned here is how traits are measured, for which measurement protocols exist I believe. I think this is a separate issue.

@@

When trait values are only available at species-level and observational data are on less precise taxonomic levels various trait aggregation methods have been used by researchers. We aggregated traits from an Australian and North American trait dataset to family-level using 5 different aggregation methods and compared the results to traits assigned at family-level for these regions. Evaluation of the differences between aggregated and assigned traits is difficult because it remains unclear what the true value of a particular trait for a particular family is. Aggregation of trait information on species or genus-level to a point estimate on family-level suggests a precision that is not necessarily present, especially not for traits with a high variability or if trait information on species-level is missing. Some traits can vary strongly on more precise taxonomic levels than the family-level. For example, Monaghan and Soares \cite{monaghan_improving_2013} found a high intra-family trait dispersion in the Tachet database for traits of the grouping features body size, flow preference, and reproduction. Furthermore, studies focusing on the lability of traits, i.e. how much traits are unconstrained by phylogeny, found that traits of ecological preferences (e.g. thermal preference), body size, resistance forms, and to a lesser extent feeding mode are labile, and thus possibly highly variable \cite{poff_functional_2006, wilkes_traitbased_2020}. However, if trait aggregation is necessary our study can give guidance, under which circumstances the choice of aggregation method might be important. The 5 aggregation methods we tested use either the mean or the median and different weightings. Our results indicated that 1) the median aggregation methods compared to mean aggregation methods were often closer to the assigned traits at family-level and 2) that the different weightings tested exerted a minor impact on the aggregated trait affinities. We obtained for both datasets more differing cases for mean aggregation methods, irrespective of the weighting. In the Australian dataset, the amount of differing cases was slightly higher for the mean aggregation methods (23 \%) than for median aggregation methods (16 to 18 \% differing cases). In the North American dataset, mean aggregation methods yielded in 47 \% of the cases a difference to the assigned traits, much higher in comparison to the median aggregation methods (between 15 and 18 \%, Table \ref{tab:summary_stat_aggr_vs_fam_assigned}). When differences occurred between median aggregation methods and assigned traits, these differences were greater compared to the mean aggregation methods, particularly for the North American dataset. There, the mean absolute differences for the median aggregations were twice as high (0.63 - 0.7) compared to the mean aggregations (0.3 - 0.31, Table \ref{tab:summary_stat_aggr_vs_fam_assigned}). The differences can be explained by the binary coding used in the North American trait dataset and in the assigned traits. As a result, traits have affinities of either 1 or 0. Therefore, traits of a particular grouping feature have in most cases a higher difference in trait affinities to each other than if they were fuzzy coded. Using the median to aggregate binary coded taxa will often result in a value of 0 or 1, while using the mean will often result in values between 0 and 1. Thus, using the median led for the North American dataset either to agreement with the assigned traits or to higher differences compared to the mean aggregation methods.

The weightings explored in this analysis should lead to different aggregated trait affinities in cases where genera within a family have different numbers of species. In fact, the simulation results indicated that for a particular family where one genus has a much larger number of species compared to the other genera, weighting the number of taxa can lead to different aggregated trait affinities, especially for traits with higher levels of variance (Figure \ref{fig:sim_indv_runs}). However, in the comparison of the trait aggregation methods to assigned traits the number of differing cases and the mean absolute differences in trait affinities were similar across the mean aggregation and across the median aggregation methods, which suggests a small influence of the weightings on the aggregation methods. Also, the distributions of absolute trait affinity differences to assigned traits were relatively similar for the mean aggregation methods and the median aggregation methods, respectively (Figure \ref{fig:diff_aggr_traits_chessman} and Figure \ref{fig:diff_aggr_traits_pyne}). The minor impact on trait aggregation of the weightings may be explained by the fact that a considerable portion of taxa had low numbers of genera or species entries. From the taxa that were compared from the North American trait dataset, 14 \% were resolved on family-level and 62 \% were resolved on genus-level. 52 \% had 5 or less genera and 13 \% contained just one genus (Figure \ref{fig:tax_hierarchy_NOA}). In the Australian dataset, 21 \% of the taxa were resolved on family-level, 40 \% were resolved on genus level, and 39 \% were resolved on species-level. 68 \% of the taxa contain 5 or less genera and 40 \% just contained one genus (Figure \ref{fig:tax_hierarchy_AUS}). Hence, these results might change when more trait information on species-level becomes available.

@@

\section*{Conclusion}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Bibliography %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\printbibliography

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{float}
\usepackage[table,xcdraw]{xcolor}
\usepackage{color, colortbl}
\usepackage{longtable}
%\usepackage[sort&compress,square,comma,authoryear]{natbib}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{rotating}
\usepackage{caption}
% \graphicspath{{/home/kunz/Dokumente/Projects/Trait_DB/Invertebrate_traits/Paper/Figures/}}
\graphicspath{{Figures/}}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{array}
\usepackage{lscape}
\usepackage{longtable}
\usepackage{hhline}
\usepackage{lmodern}
\usepackage[
backend = biber,
style = apa,
citestyle = authoryear,
maxcitenames = 2, 
maxbibnames = 99 % ,
%sorting = none % sort by name year title
]{biblatex}
\addbibresource{Ref_invertebrate_DB.bib}

% From freshwater biology:
% List all sources in the reference list alphabetically by name. In text citations should follow the author-date method. This means that the author's last name and the year of publication for the source should appear in the text, for example, (Jones, 1998), and a complete reference should appear in the reference list at the end of the paper.

% References are styled according to the sixth edition of the Publication Manual of the American Psychological Association. A sample of the most common entries in reference lists appears below. Please note that for journal articles, issue numbers are not included unless each issue in the volume begins with page one.

% Article: 
% One author: Fawcett, T. (2006). An introduction to ROC analysis. Pattern Recognition Letters, 27(8), 861–874. DOI: 10.1016/j.patrec.2005.10.010.

% 2 to 7 authors: Daley, C. E., & Nagle, R. J. (1996). Relevance of WISC-III Indicators for assessment of learning disabilities. Journal of Psychoeducational Assessment, 14 (4), 320–333.

% More than 7 authors: Rutter, M., Caspi, A., Fergusson, D., Horwood, L. J., Goodman, R., Maughan, B., … Carroll, J. (2004). Sex differences in developmental reading disability: New findings from 4 epidemiological studies. Journal of the American Medical Association, 291(16), 2007–2012. DOI: 10.1001/jama.291.16.2007

% In press or forthcoming: van Bergen, E., de Jong, P. F., Maassen, B., Krikhaar, E., Plakas, A., & van der Leij, A. (in press). IQ of four-year-olds who go on to develop dyslexia. Journal of Learning Disabilities. DOI: 10.1177/0022219413479673

@@

Explaining and predicting how communities are shaped by environmental factors is one of the main goals of ecology. Species traits, defined as measurable properties of an organism (\cite{mcgill_rebuilding_2006}), might be beneficial in achieving this goal (\cite{heino_jani_macroecological_2013}). Traits evolve through adaptations (e.g., physiological, behavioural, etc.) of organisms to their environment and indicate direct or indirect linkages between the biological response of an organism or a population and its environment (\cite{southwood_habitat_1977, verberk_delivering_2013}).
Besides providing a mechanistic explanation of species-environment relationships, trait-based approaches might be suitable for large scale analysis because the variability in trait responses is lower than in taxonomic responses (\cite{bonada_taxonomic_2007, baird_toward_2011}). Traits of freshwater invertebrate individuals are difficult to determine because - unlike plant traits - they often can not be measured directly. For example, to gain knowledge on feeding habits requires evaluating mouthpart morphology, consumed food, and the organisms function within its community (\cite{moog_comprehensive_nodate}). Nevertheless, invertebrate traits have been increasingly used in freshwater ecology, e.g. by relating macroinvertebrate trait composition to environmental factors and as trait metrics in biomonitoring (\cite{poff_developing_2010, szocs_effects_2014, bhowmik_large_2015, menezes_beyond_2010}).

In the last decades, freshwater ecologists have compiled comprehensive invertebrate trait databases for various continents (\cite{usseglio-polatera_biomonitoring_2000, schmidt-kloiber_www.freshwaterecology.info_2015, vieira_database_nodate, Philips_and_Smith_NZ_DB_2018, kefford_integrated_2020, tomanova_trophic_2006}). The availability of invertebrate trait data from different continents enables comparisons of trait variation and analyses of their relation to environmental factors across large scales. However, such analyses have been carried out mostly within continents, using information from one or two trait databases. For example, Bonada et al. (\cite{bonada_taxonomic_2007}) compared trait composition for Mediterranean and temperate regions in Europe using traits from Usseglio-Polatera et al. (\cite{usseglio-polatera_biomonitoring_2000}, typically referred to as Tachet database); Poff et al. \cite{poff_developing_2010} characterized trait composition across sites in the Western US using traits from Poff et al. (\cite{poff_functional_2006}), and Botwe et al. \cite{botwe_effects_2018} investigated the effect of salinity on invertebrate traits in different sites in South Australia using trait data from Poff et al. (\cite{poff_functional_2006}) and Schäfer et al. (\cite{schafer_trait_2011}). Analyses of invertebrate traits that synthesize information on invertebrate grouping features from more than two different continents are rare.

In this study, we follow the terminology proposed by Schmera et al. (\cite{schmera_proposed_2015}), where a grouping feature is defined as a general property (e.g. feeding mode) that comprises a "group of related traits (e.g., predator, shredder, etc.) that vary among species or among individuals within a species". Thus, we use the term grouping feature in place of which many studies use the term "trait", and the term trait instead of "trait state", "modality" or "trait category". Traits can be described using different codings (binary, fuzzy) that represent the prevalence of a characteristic in an organism.

To our knowledge, only Brown et al. (\cite{brown_functional_2018}) harmonised grouping features from more than two geographically distant invertebrate trait databases, for a limited set of grouping features (8) and taxa (112), in a study on the influence of decreasing glacier cover on functional diversity and community assembly of invertebrates. We suspect that the heterogeneity of information in freshwater invertebrate trait databases, besides the diversity of taxa across regions, is likely a major reason for the lack of studies across continents. To harmonize grouping features from different regions, first commonly accepted and unambiguous trait definitions are required (\cite{schneider_towards_2019}). In the best case, grouping features would be classified into the same traits across databases or they could easily be harmonised using standardised terminology. However, a lack of standardised terminology of trait definitions and poor metadata quality in many trait databases are common issues throughout the field of trait-based ecology (\cite{baird_toward_2011, kissling_towards_2018}). Secondly, consistent coding of traits facilitates the compatibility of trait data from different databases.

@@

However, invertebrate trait databases are heterogeneous regarding the coding they use for their traits (\cite{culp_incorporating_2011}). For example, in the study from Brown et al. (\cite{brown_functional_2018}) the authors needed to reclassify traits using expert knowledge given that the individual databases employed different coding approaches (i.e. European and New Zealand databases fuzzy coding, North American database binary coding). Databases with standardised and unambiguous traits and a consistent coding of traits would minimize the data processing effort. Discrepancies in the taxonomic resolutions (e.g. species, genus or family) when linking observational taxonomic data to trait database represents another challenge. When observations are on a more precise taxonomic level than data available in the trait databases (e.g. observations on species-level, trait data on genus-level) trait information of the less precise taxonomic level is often assigned (e.g. \cite{szocs_effects_2014, vos_taxonomic_2017}). Conversely, if trait information is only available on more precise taxonomic levels than the observed taxa, traits are aggregated to a less precise taxonomic level (e.g. \cite{poff_functional_2006, szocs_effects_2014, piliere_a._f._h._importance_2016, aspin_extreme_2019}). Studies have used different methods of trait aggregation, e.g. the mean (\cite{magliozzi_functional_2019}), median (\cite{szocs_effects_2014}) or the mode (\cite{piliere_a._f._h._importance_2016}). Up to now, studies on how and to which extent different trait aggregation methods influence trait-based analysis are missing. However, related knowledge would inform future studies regarding the choice of the aggregation method.

The outlined difficulties when working with or synthesizing trait data motivated us to (1) investigate discrepancies in trait definition between trait databases from Europe, North America, Australia, and New Zealand. As a basis for further evaluating different trait aggregation methods and how harmonising definition discrepancies between grouping features can influence trait-environment relationships, we established 4 novel invertebrate trait datasets from these trait databases by harmonising 7 grouping features. We evaluate different trait aggregation methods by (2) comparing aggregated traits with assigned traits by experts at family-level for two of the established trait datasets. (3) we re-analysed data on the effect of anthropogenic salinisation on biological traits by Szöcs et al. (\cite{szocs_effects_2014}) using trait data from the established harmonised European trait datasets and aggregated traits. By comparison with the original analysis, we investigated how harmonising and aggregating trait data can modify the outcome of trait-environment relationships.

@@

Using harmonised grouping features resulted in less traits that distinguish upstream and downstream sites in comparison to the original analysis (Figure \ref{fig:violin_plot_species_sc} and Figure \ref{fig:boxplots_scores_on_constrained_axis}). According to the RDA of the trait composition, downstream sites were characterised by taxa with the traits multivoltinism, ovoviviparity and long life cycle duration. Upstream sites were characterised by univoltine taxa with a short life cycle duration that lay their eggs in an aquatic environment (aquatic eggs). The trait aquatic eggs of the harmonised grouping feature oviposition has been derived by amalgamating the trait oviposition in clutches and other related traits (Table \ref{tab:traits_harmonisation}). The traits shredder and gill respiration did not characterise sites with high salinisation.

Using at family-level aggregated traits from the harmonised dataset showed also results similar to the original analysis but with fewer traits distinguishing upstream and downstream sites (Figure \ref{fig:violin_plot_species_sc} and Figure \ref{fig:boxplots_scores_on_constrained_axis}). The \textit{direct\_agg\textsubscript{mean}}, \textit{direct\_agg\textsubscript{median}}, and \textit{weighted\_agg} 
characterised the downstream sites with the same traits as the original analysis except that downstream sites were not characterised by the trait shredder. Upstream sites were characterised by the traits univoltinism, aquatic eggs and short life cycle duration. The same results were obtained when re-analysing the data with traits aggregated by the \textit{stepwise\_agg\textsubscript{mean}} and \textit{stepwise\_agg\textsubscript{median}}, with the exception that none of the life cycle traits characterised upstream or downstream sites. Thus, in our comparison aggregating by \textit{direct\_agg\textsubscript{mean}}, \textit{direct\_agg\textsubscript{median}}, and \textit{weighted\_agg} yielded to the least change in interpretation of the RDA results compared to the original results.

\begin{figure}[H]
    \centering
    \includegraphics[width=16.5cm, height=10cm]{Species_scores_rda.png}
    \caption{Species scores obtained by RDA from the original analysis \cite{szocs_effects_2014}, using harmonised grouping features, and using harmonised grouping features with traits aggregated to family-level. The dots represent the individual species scores for each analysed trait along the conductivity axis. The violin plot shows the density estimate of the species scores. Gray vertical lines indicate the median of the obtained species scores.}
    \label{fig:violin_plot_species_sc}
\end{figure}

% Old figure:
% \begin{sidewaysfigure}
%   \label{fig:boxplots_scores_on_constrained_axis}
%   \centering
%   \includegraphics[width= 22cm, height=12cm]{Scores_on_constrained_axis_combined.pdf}
%   \caption{RDA of traits constrained by electric conductivity for the tested methods and the original study. Shown are boxplots of the site scores along the conductivity axis. The rug on the right side of each plot indicates species scores of the traits on the conductivity axis. Only traits with a mahalanobis distance greater than the $97.5 \%$ quantile of the Chi-square distribution (5.02) were labelled.}
% \end{sidewaysfigure}

\begin{figure}[H]
  \centering
  \includegraphics[width=16.5cm, height=25cm]{Scores_on_constrained_axis_comb_ALT.pdf}
\end{figure}
\clearpage % insert a page break
\captionof{figure}{RDA of traits constrained by electric conductivity for the tested methods and the original study. Shown are boxplots of the site scores along the conductivity axis. The rug on the right side of each plot indicates species scores of the traits on the conductivity axis. Only traits with a mahalanobis distance greater than the $97.5 \%$ quantile of the Chi-square distribution (5.02) were labelled. For better comparability, species scores of the original analysis were multiplied by -1.} 
\label{fig:boxplots_scores_on_constrained_axis}

@@

% From Freshwater biology: 
% This should highlight the significance of the results and place them in the context of other work. It should not include references to figures or tables cited in the Results – check that all of your Results are presented in the Results section. Figures (e.g. photographs, conceptual diagrams) or tables presented specifically as part of the Discussion are welcome and may be cited there.

@@

Synthesizing trait information from multiple trait databases is crucial to unfold the full potential of trait-based approaches, e.g. for studying community trait responses to environmental gradients worldwide. Our attempt of harmonising invertebrate trait information from trait databases of different regions showed that grouping feature harmonisation is labour-intensive because grouping features are differentiated differently into traits between trait databases and various codings are used to describe traits. In addition, the same traits are sometimes defined differently in the investigated databases, requiring expert knowledge to reclassify these traits (e.g. the trait piercer in the Tachet database). Although, others have noted the lack of standardised trait terminology in freshwater ecology \cite{baird_toward_2011, brink_traits-based_2011}, we are the first to describe invertebrate trait definition discrepancies for some commonly used grouping features.
%In addition we provide harmonised grouping feature trait datasets that can be used for other studies
To resolve definition discrepancies and facilitate data synthesis in the future, terminological standards are needed. Harmonised definitions and concepts of traits have been developed in the past for other organism groups, e.g. for plants with the \textit{Thesaurus Of Plant characteristics} (TOP) initiative \cite{garnier_towards_2017}. The core of this harmonisation initiative is to provide standardised trait definitions and to draw connections to synonyms, related terms, and surrounding concepts by linking to other controlled vocabularies or ontologies in the field. Such initiatives could be a role model for freshwater ecologists to establish unambiguous terminologies for invertebrate traits. The existing freshwater invertebrate trait databases could be linked through standardised terminologies or ontologies, as suggested by Baird et al. \cite{baird_toward_2011} almost a decade ago. By following the recently proposed \textit{Ecological Trait-data Standard Vocabulary} (ETS) providers of invertebrate trait data could connect their traits to such ontologies \cite{schneider_towards_2019}. Once standardised trait definitions are established, these will improve trait data sharing, trait data processing when working with multiple trait databases, and ultimately interpretation of the derived results.

@@

When trait values are only available at species-level and observational data are on less precise taxonomic levels various trait aggregation methods have been used by researchers. We aggregated traits from an Australian and North American trait dataset to family-level using 5 different aggregation methods and compared the results to traits assigned at family-level for these regions. Evaluation of the differences between aggregated and assigned traits is difficult because it remains unclear what the true value of a particular trait for a particular family is. Aggregation of trait information on species or genus-level to a point estimate on family-level suggests a precision that is not necessarily present, especially not for traits with a high variability or if trait information on species-level is missing. Some traits can vary strongly on more precise taxonomic levels than the family-level. For example, Monaghan and Soares \cite{monaghan_improving_2013} found a high intra-family trait dispersion in the Tachet database for traits of the grouping features body size, flow preference, and reproduction. Furthermore, studies focusing on the lability of traits, i.e. how much traits are unconstrained by phylogeny, found that traits of ecological preferences (e.g. thermal preference), body size, resistance forms, and to a lesser extent feeding mode are labile, and thus possibly highly variable \cite{poff_functional_2006, wilkes_traitbased_2020}. However, if trait aggregation is necessary our study can give guidance, under which circumstances the choice of aggregation method might be important. The 5 aggregation methods we tested use either the mean or the median and different weightings. Our results indicated that 1) the median aggregation methods compared to mean aggregation methods were often closer to the assigned traits at family-level and 2) that the different weightings tested exerted a minor impact on the aggregated trait affinities. We obtained for both datasets more differing cases for mean aggregation methods, irrespective of the weighting. In the Australian dataset, the amount of differing cases was slightly higher for the mean aggregation methods (23 \%) than for median aggregation methods (16 to 18 \% differing cases). In the North American dataset, mean aggregation methods yielded in 47 \% of the cases a difference to the assigned traits, much higher in comparison to the median aggregation methods (between 15 and 18 \%). When differences occurred between median aggregation methods and assigned traits, these differences were greater compared to the mean aggregation methods, particularly for the North American dataset. There, the mean absolute differences for the median aggregations were twice as high (0.63 - 0.7) compared to the mean aggregations (0.3 - 0.31). The differences can be explained by the binary coding used in the North American trait dataset and in the assigned traits. As a result, traits have affinities of either 1 or 0. Therefore, traits of a particular grouping feature have in most cases a higher difference in trait affinities to each other than if they were fuzzy coded. Using the median to aggregate binary coded taxa will often result in a value of 0 or 1, while using the mean will often result in values between 0 and 1. Thus, using the median led for the North American dataset either to agreement with the assigned traits or to higher differences compared to the mean aggregation methods.

The weightings explored in this analysis should lead to different aggregated trait affinities in cases where genera within a family have different numbers of species. In fact, the simulation results indicated that for a particular family where one genus has a much larger number of species compared to the other genera, weighting the number of taxa can lead to different aggregated trait affinities, especially for traits with higher levels of variance. However, in the comparison of the trait aggregation methods to assigned traits the number of differing cases and the mean absolute differences in trait affinities were similar across the mean aggregation and across the median aggregation methods, which suggests a small influence of the weightings on the aggregation methods. Also, the distributions of absolute trait affinity differences to assigned traits were relatively similar for the mean aggregation methods and the median aggregation methods, respectively. The minor impact on trait aggregation of the weightings may be explained by the fact that a considerable portion of taxa had low numbers of genera or species entries. From the taxa that were compared from the North American trait dataset, 14 \% were resolved on family-level and 62 \% were resolved on genus-level. 52 \% had 5 or less genera and 13 \% contained just one genus (Figure \ref{fig:tax_hierarchy_NOA}). In the Australian dataset, 21 \% of the taxa were resolved on family-level, 40 \% were resolved on genus level, and 39 \% were resolved on species-level. 68 \% of the taxa contain 5 or less genera and 40 \% just contained one genus (Figure \ref{fig:tax_hierarchy_AUS}). Hence, these results might change when more trait information on species-level becomes available.

@@

\subsection*{Outlook}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Bibliography %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\printbibliography
